On the problem of repeated supervised learning

### 2023-03-08 paper/main.tex
Add files via upload

    In this paper, we explore some aspects of continuous learning artificial intelligence systems as they interact with and influence their environment. For the process of repeated and multiple learning, prediction and updating of the sample, we built a mathematical model and investigated the properties of this process based on the constructed model. We consider this problem in a novel (as far as we know) sense: from the point of view of functional analysis and probability theory. Based on our results we will make several synthetic experiments.

@@

    In this paper we consider the problem of repeated supervised learning \begin{otherlanguage}{russian}(многократное машинное обучение)\end{otherlanguage}, in which the training sample is not fixed, but is updated depending on the predictions of the trained model on the test sample \cite{ma2020machine, khritankov2021existence, jiang2019degenerate}. This problem is widespread in many areas of machine learning and artificial intelligence, but there is still no complete theoretical description of this method of model learning, and we have tried to correct this in this paper. For example repeated supervised learning appears in recommendation systems \cite{khritankov2021existence, sinha2016deconvolving}, healthcare \cite{adam2020hidden} and predictive policing \cite{ensign2018runaway}. Our main goal was to consider the problem of multiple machine learning from a theoretical point of view, using probability theory, statistics and functional analysis.

@@

    Main contributions of this paper are as follows. In Section \ref{Related_work} we will make a literature review of our problem (which will not be particularly large, since the problem we are considering is quite new). In Section \ref{Problem_statement} we will build a mathematical model for the process of multiple learning, prediction and updating of the sample and outline the main questions, the answers to which we will explore in Section \ref{Main_results}. In Section \ref{Main_results} we will formulate several theorems (sufficient conditions) for the mapping $\text{D}$, which are novel to the literature (as far as we know). In Section \ref{Experiments}, based on the results from Section \ref{Main_results} we will provide some synthetic experiments.

@@

    Also our problem is connected with dynamical systems, which are studied in \cite{katok1995introduction, nemytskii2015qualitative}, but all these books consider continuous time, and in our work it should be discrete.

    Some authors analyzed Markov processes from the point of view of dynamical systems \cite{tarlowski2017global, vershik2005does}, but in Markov chain the future of the process does not depend on the past, but in repeated supervised learning that's not so. Other authors \cite{varvenne2019rate, pap1996fixed} studied stochastic dynamics systems in general.

    All theoretical results on probability theory were taken from the book \cite{shiryaev2016probability}, on functional analysis the main sources were \cite{kolmogorov1957elements, rudin1991functional}, the results on the theory of measure and the Lebesgue integral were taken from \cite{rudin2006real}.

@@

    Let's consider an autonomous (time-independent explicitly) discrete dynamical system (there is a dedicated variable - the step number, which increases), at step $t$ and $t+1$ of which the ratio is fulfilled

    \begin{equation*}
        f_{t+1}(x) = \text{D}(f_t)(x), \quad \text{for}~ \forall x \in \mathbb{R}^n
    \end{equation*}

@@

    \begin{enumerate}
        \item[\textbf{1.}]  What are the conditions so that the mapping $\text{D}$ becomes a transformation?

@@

>>>>

    In this research paper, we delve into the intricacies of continuous learning artificial intelligence systems as they interact with and influence their environment. We develop a mathematical model to examine the process of repeated and multiple learning, prediction, and dataset updating. Our investigation of this process is based on the principles of functional analysis and probability theory, which is a novel approach to this problem. We aim to conduct several synthetic experiments based on our findings, hoping to contribute to a better understanding of the behavior of continuous learning AI systems.

@@

    In this paper we consider the problem of repeated supervised learning \begin{otherlanguage}{russian}(многократное машинное обучение)\end{otherlanguage}, in which the training sample is not fixed, but is updated depending on the predictions of the trained model on the test sample \cite{ma2020machine, khritankov2021existence, jiang2019degenerate}. In many applications, the use of multiple learning techniques can actually lead to suboptimal results. Our main goal was to present a mathematical theory that explains why combining multiple learning algorithms can sometimes hinder their effectiveness. Repeated supervised learning appears in many machine learning applications, for example in recommendation systems \cite{khritankov2021existence, sinha2016deconvolving}, healthcare \cite{adam2020hidden} and predictive policing \cite{ensign2018runaway}.

@@

    Main contributions ...

    Structure of this paper are as follows. In Section \ref{Related_work} we compare our article with the works of other authors and show its novelty to the literature. In Section \ref{Problem_statement} we build a mathematical model for the process of multiple learning, prediction and updating of the sample and outline the main questions, the answers to which we explore in Section \ref{Main_results}. In Section \ref{Main_results} we provide our main contributions for the mapping $\text{D}$. In Section \ref{Experiments}, based on the results from Section \ref{Main_results}, we conduct some synthetic experiments.

@@

    Also our problem is connected with dynamical systems, which are studied in \cite{katok1995introduction, nemytskii2015qualitative}, but all these works consider continuous time, and in our work it should be discrete.

    Some authors analyzed Markov processes from the point of view of dynamical systems \cite{tarlowski2017global, vershik2005does}. But in Markov chain the future of the process does not depend on the past. Other authors \cite{varvenne2019rate, pap1996fixed} studied stochastic dynamics systems in general.

    An important contribution of this paper is that ...

@@

    Let's consider an autonomous (time-independent explicitly) discrete dynamical system. There is a dedicated variable - the step number, which increases, at step $t$ and $t+1$ of which the ratio is fulfilled

    \begin{equation*}
        f_{t+1}(x) = \text{D}(f_t)(x), \quad \text{for}~ \forall x \in \mathbb{R}^n,
    \end{equation*}

@@

    \begin{enumerate}
        \item[\textbf{1.}]  What are the conditions so that the mapping $\text{D}$ becomes a transformation $\mathbf{R}$ into $\mathbf{R}$?

@@

    Let's analyze the importance of each question. We consider $\text{D}$ as operator of changing distribution of our dataset, so it has to transform $\mathbf{R}$ into $\mathbf{R}$. If $\text{D}$ is a contraction mapping, then for any starting function $f_0$ after a sufficiently large T, the result of our algorithm after $T$ steps would be $D^T(f_0) = f^*$, where $f^*$ is a fixed point, so after $T$ steps metrics of our algorithm cannot be improved. If $\text{D}$ has a fixed point $f$, when if $f_0 = f$ and we cannot improve our metrics as well.
